%%% The main file. It contains definitions of basic parameters and includes all other parts.

%% Settings for single-side (simplex) printing
% Margins: left 40mm, right 25mm, top and bottom 25mm
% (but beware, LaTeX adds 1in implicitly)
\documentclass[12pt,a4paper]{report}
\setlength\textwidth{145mm}
\setlength\textheight{247mm}
\setlength\oddsidemargin{15mm}
\setlength\evensidemargin{15mm}
\setlength\topmargin{0mm}
\setlength\headsep{0mm}
\setlength\headheight{0mm}
% \openright makes the following text appear on a right-hand page
\let\openright=\clearpage

%% Settings for two-sided (duplex) printing
% \documentclass[12pt,a4paper,twoside,openright]{report}
% \setlength\textwidth{145mm}
% \setlength\textheight{247mm}
% \setlength\oddsidemargin{14.2mm}
% \setlength\evensidemargin{0mm}
% \setlength\topmargin{0mm}
% \setlength\headsep{0mm}
% \setlength\headheight{0mm}
% \let\openright=\cleardoublepage

%% Character encoding: usually latin2, cp1250 or utf8:
\usepackage[utf8]{inputenc}

%% Further useful packages (included in most LaTeX distributions)
\usepackage{amsmath}        % extensions for typesetting of math
\usepackage{amsfonts}       % math fonts
\usepackage{amsthm}         % theorems, definitions, etc.
\usepackage{bbding}         % various symbols (squares, asterisks, scissors, ...)
\usepackage{bm}             % boldface symbols (\bm)
\usepackage{graphicx}       % embedding of pictures
\usepackage{fancyvrb}       % improved verbatim environment
\usepackage{natbib}         % citation style AUTHOR (YEAR), or AUTHOR [NUMBER]
\usepackage[nottoc]{tocbibind} % makes sure that bibliography and the lists
			    % of figures/tables are included in the table
			    % of contents
\usepackage{dcolumn}        % improved alignment of table columns
\usepackage{booktabs}       % improved horizontal lines in tables
\usepackage{paralist}       % improved enumerate and itemize
\usepackage[usenames]{xcolor}  % typesetting in color
\usepackage{url}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{booktabs, caption, fixltx2e}
\usepackage[flushleft]{threeparttable}
\usepackage[scaled=.95]{inconsolata}
\usepackage{listings}
\usepackage{tikz, tikz-qtree}
\usepackage{float}
\usepackage{svg}
\usepackage{xspace}

\newcommand*{\incon}[1]{{\fontfamily{fi4}\selectfont #1}}
\def\cpp{{C\nolinebreak[4]\hspace{-.05em}\raisebox{.4ex}{\tiny\bf ++\xspace}}}

\lstset{language=[5.2]Lua, basicstyle=\ttfamily, columns=fullflexible, showstringspaces=false, commentstyle=\color{gray}}

% Listing float:
\newfloat{listing}{tbhp}{lst}
\floatname{listing}{Listing}

%%% Basic information on the thesis

% Thesis title in English (exactly as in the formal assignment)
\def\ThesisTitle{The Dungeon Throne: A 3D Dungeon Management Game}

% Author of the thesis
\def\ThesisAuthor{Jaroslav Jindrák}

% Year when the thesis is submitted
\def\YearSubmitted{2016}

% Name of the department or institute, where the work was officially assigned
% (according to the Organizational Structure of MFF UK in English,
% or a full name of a department outside MFF)
\def\Department{Department of Distributed and Dependable Systems}

% Is it a department (katedra), or an institute (ústav)?
\def\DeptType{Department}

% Thesis supervisor: name, surname and titles
\def\Supervisor{Mgr. Pavel Ježek, Ph.D.}

% Supervisor's department (again according to Organizational structure of MFF)
\def\SupervisorsDepartment{Department of Distributed and Dependable Systems}

% Study programme and specialization
\def\StudyProgramme{Computer~Science}
\def\StudyBranch{Programming and Software Systems}

% An optional dedication: you can thank whomever you wish (your supervisor,
% consultant, a person who lent the software, etc.)
\def\Dedication{%
I would like to thank my supervisor, Pavel Ježek, for his patience and advice that he has provided to me.
Without his guidance this thesis would hardly reach the level of quality I believe it to have.

I would also like to thank my parents, who supported me throughout my work on this thesis and without whom I would
never be able to dedicate the time needed to finnish this project.
}

% Abstract (recommended length around 80-200 words; this is not a copy of your thesis assignment!)
\def\Abstract{%
    Dungeon Keeper is an iconic and genre-defining game that places the player in the role of a dungeon master that
    has to protect their dungeon against groups of heroic enemies that try to steal the treasures located in the
    dungeon. However, it does not provide any tools that would allow user created modifications and thus the replayability
    of the game might not reach its full potential.

    In this thesis, we examine the dungeon management genre and the aspects of game modifiability and use the knowledge gained
    from this examination to design a spiritual successor to Dungeon Keeper that offers tools that allow its players
    to create and distribute modifications that can be applied to the game.

    We then investigate different tools and approaches that can be used for game development and use those we find
    to be suitable to implement our game. The game we created is true to the play style of the original Dungeon Keeper
    and provides an API that can be used in the Lua programming language to create modifications of the game's elements.
}

% 3 to 5 keywords (recommended), each enclosed in curly braces
\def\Keywords{%
    {dungeon management game}, {Lua scripting}, {Ogre}, {CEGUI}
}

%% The hyperref package for clickable links in PDF and also for storing
%% metadata to PDF (including the table of contents).
\usepackage[pdftex,unicode]{hyperref}   % Must follow all other packages
\hypersetup{breaklinks=true}
\hypersetup{pdftitle={\ThesisTitle}}
\hypersetup{pdfauthor={\ThesisAuthor}}
\hypersetup{pdfkeywords=\Keywords}
\hypersetup{urlcolor=blue}

% Definitions of macros (see description inside)
\include{macros}

% Title page and various mandatory informational pages
\begin{document}
\include{title}

%%% A page with automatically generated table of contents of the bachelor thesis

\tableofcontents

%%% Each chapter is kept in a separate file
\include{chap01}
\include{chap02}
\include{chap03}
\include{chap04}
\include{chap05}

\include{epilog}

%%% Bibliography
\include{bibliography}

%%% Attachments to the bachelor thesis, if any. Each attachment must be
%%% referred to at least once from the text of the thesis. Attachments
%%% are numbered.
%%%
%%% The printed version should preferably contain attachments, which can be
%%% read (additional tables and charts, supplementary text, examples of
%%% program output, etc.). The electronic version is more suited for attachments
%%% which will likely be used in an electronic form rather than read (program
%%% source code, data files, interactive charts, etc.). Electronic attachments
%%% should be uploaded to SIS and optionally also included in the thesis on a~CD/DVD.
\chapwithtoc{Attachments}

Attachments, which can be found on the attached DVD contain:

\begin{enumerate}[label=\textbf{\Alph*.}]
    \item Implementation of the game, which we can find in directory \texttt{tdt-project} and which contains source code, project files,
        compiled libraries and resources the game uses. It contains the following files and directories:
        \begin{itemize}
            \item Directory \texttt{bin} in which the Visual Studio solution generates the output of the compilation.
            \item Directory \texttt{deps}, which contains the Visual C++ Redistributable that is needed by the game.
            \item Directory \texttt{lib}, which contains \texttt{.dll} and \texttt{.lib} files of Ogre3D, CEGUI and Lua as well as their
                header files. These libraries were compiled from the official source code and were  included because there are no official
                precompiled packages of these libraries for Visual Studio 2015.
            \item Directory \texttt{res}, which contains the directory \texttt{ogre-configs} -- the contents of which need to be placed
                in the same directory as the game's executable file -- and the directory \texttt{resources}, which needs to be placed
                in the same directory as the game's executable file.
            \item Directory \texttt{src}, which contains the source code of our game.
            \item Visual Studio 2015 solution files, namely \texttt{tdt-project.sln} which represents the solution,
                \texttt{tdt-project.vcxproj} which represents the project and \texttt{tdt-project.vcxproj.filters} which represents
                the source code hierarchy in the project.
        \end{itemize}
    \item Documentation of the game and of the modding API, which we can find in directory \texttt{docs}. It contains the following files
        and directories:
        \begin{itemize}
            \item File \texttt{components.txt}, which contains a list of all components and information on how to include these
                in a table that defines an entity.
            \item Directory \texttt{lua-api}, which contains files that represent different modules of the Lua API.
            \item Directory \texttt{engine-doxy}, which contains HTML and PDF documentation of the \cpp engine generated by
                Doxygen~\cite{doxy}.
        \end{itemize}
    \item Compiled version of the game, which we can find in directory \texttt{tdt-game}. It contains the game's executable file
        called \texttt{tdt.exe} and all files the game needs in order to run -- resources, configurational files, scripts and libraries.
    \item Electronic version of this thesis.
\end{enumerate}

\openright
\end{document}
